---
title: "Affordable Learning and Financial Support"
editor: visual
---

This is a chapter about affordable learning and financial support at University of California San Diego

```{r}
#| label: run_scripts
#| include: FALSE

source("rscripts/load_packages.R", local = knitr::knit_global())

source("rscripts/functions_clean_data.R" ,local = knitr::knit_global())

source("rscripts/functions_visualize.R" ,local = knitr::knit_global())

```

```{r}
#| label: load_data
#| include: FALSE

#ucues
es<- read_excel("data/ucues_2022_rebuild.xlsx")%>%
  filter(wg=="Affordable")%>%
  label_groups(.)

ucall<-read_excel("data/ucues_alluc.xlsx")

gses<-read_excel("data/UCGSES.xlsx")%>%
  filter(wg=="Affordable")%>%
  label_groups(.)

ucall_gses<-read_excel("data/gses_alluc.xlsx")

```

```{r}
#| label: viz_captions
#| include: FALSE

#Diff labels
lab_diff_axis_ug<-"Difference from all undergraduate average"
lab_diff_axis_grad<-"Difference from all graduate student average"

#subtitle
subtitle_grad_agree<-"UCSD Graduate Students who somewhat agree, agree, or strongly agree"

#caption
caption_ug<-"Source: UCUES 2022, UC Infocenter"
caption_grad<-"Source: GSES 2021 UC Infocenter"
```

## Undergraduates

All data come from UCUES

### Food Security

Student respondents are assigned a food security level: very low food security, low food security, or food secure, based on their responses to six questions.

For more information on this measure, please visit [UCOP Measuring Food Insecurity](https://www.universityofcalifornia.edu/sites/default/files/measuring-food-insecurity.pdf)

#### Food Security over Time

These visuals display the percentage respondents who are "food secure. For example, in 2022 58% of UCSD undergraduate respondents were classified as food secure. This means the remaining 42% of respondents were food insecure- either low food security or very low food security.

The percent of undergraduates who are food secure has remained relatively stable over time -between roughly 50-60%. UCSD's food security rates are slightly higher or equal to the overal UC system rate.

```{r}
#| label: food_secure_ug_overtime
#| echo: FALSE
#| warning: FALSE

foodsecure_uccomp<-ucall%>%
  filter(q=="Food Security"&campus=="UC")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="positive")%>%
  ungroup(.)%>%
  mutate(campus="All UC")

food_secure_time<-es%>%
  #select question and comparison group - leave all years
  filter(q=="Food Security"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="positive")%>%
  ungroup(.)%>%
  mutate(campus="UCSD")%>%
  bind_rows(., foodsecure_uccomp)

  
  g_foodsecure_time<-plot_time(food_secure_time)+
  #need to tweak scale because otherwise it's too zoomed in
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(.5,.75))+
   labs(title="Food Secure",
        caption=caption_ug,
      subtitle = 'Percent of Food Secure respondents',
         x="UCUES Survey Year")

direct.label(g_foodsecure_time,list(cex=30,  "first.qp"))

```

#### Food Security Differences by Group

Focusing on UCSD students in 2022, we find that there are significant differences by student group. Students who are not First Generation (64%), not Pell Grant Recipients (64%), or are White (63%) or Asian (61%) as defined by UCOP have higher rates of food security than the student body average of 58%. Food security among international students is the same as the student body average.

Conversely, the food security rates of American Indian students (50%), First Generation students (49%) Pell Grant recipients (48%), Hispanic or Latinx students (48%) or African American students(44%), are less than the student body average. For all these groups, at least half of students are food insecure. The span of the difference is 20 percentage points- between students who don't recieve Pell Grants (64%) and African American students (44%).

```{r}
#| label: food_secure_ug_differences_2022
#| echo: FALSE
#| warning: FALSE
foodsecure_all_2022<-es%>%
filter(q=="Food Security"&year=="2022"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
collapse_positive_vibe(.)

foodsecure_2022<-es%>%
  filter(q=="Food Security"&year=="2022")%>%
  clean_vibe_foodsecure(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=foodsecure_all_2022$value)%>%
  create_diff_cols(.)

  
plot_lollipop_div(foodsecure_2022)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.15,.1),
  #second axis- real value
  sec.axis = sec_axis(~ . +foodsecure_all_2022$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="Food secure",
       subtitle ="UCSD Undergrads who are food secure",
       y=lab_diff_axis_ug,
       caption=caption_ug)
```

### Housing Costs

The prioritization poll selected two housing questions from the most recent UCUES 2022 wave- these questions had not been asked previously. There are other housing related questions future versions of this document can add to provide historical context.

The question block prompts: "For the following statements, please indicate whether the statement was never true, sometimes true, or often true for you during the current academic year:"

"I was unable to pay all the cost of my housing on time."

"I was worried I would not have enough money to cover the cost of my housing"

Respondents could select "Never True", "Sometime True" and "Often True." For consistency in selecting positive responses - I focus on "Never True" responses- meaning that respondents never were unable to pay (always able to pay) and never worried about their housing costs.

```{r}
#| label: unable_pay_house_ug_diff_2022
#| echo: FALSE
#| warning: FALSE
payhousingall_2022<-es%>%
filter(q=="I was unable to pay all of the cost of my housing on time."&year=="2022"&comp=="All")%>%
  clean_vibe_true3(.)%>%
collapse_positive_vibe(.)



payhousing_2022<-es%>%
  filter(q=="I was unable to pay all of the cost of my housing on time."&year=="2022")%>%
   clean_vibe_true3(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=payhousingall_2022$value)%>%
  create_diff_cols(.)

  
plot_lollipop_div(payhousing_2022)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.15,.15),
  #second axis- real value
  sec.axis = sec_axis(~ . +payhousingall_2022$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I was always able to pay all the cost \n of my housing on time",
       subtitle ="UCSD Undergrads who could always pay all housing cost on time",
       y=lab_diff_axis_ug,
       caption=caption_ug)
```

About 80% of students were always able to pay for housing costs on time at UCSD, meaning 20% of students were sometimes or often unable to pay on time. THis propotion is slightly higher than the UC average of 77% alway able to pay (not pictured). This division differs by students groups. Proportions differ by 17 percentage points, from 72% for African American students to 89% for American Indian students.

However, most students worry about paying for housing. Overall 46% never worry -similar to the UC average of 45%.

```{r}
#| label: worried_pay_house_ug_diff_2022
#| echo: FALSE
#| warning: FALSE


worriedhousingall_2022<-es%>%
filter(q=="I worried I would not have enough money to cover the cost of my housing." & year=="2022"& comp==
"All")%>%
  clean_vibe_true3(.)%>%
collapse_positive_vibe(.)

worriedhousing_2022<-es%>%
  filter(q=="I worried I would not have enough money to cover the cost of my housing." & year=="2022")%>%
   clean_vibe_true3(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=worriedhousingall_2022$value)%>%
  create_diff_cols(.)

  
plot_lollipop_div(worriedhousing_2022)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.17,.15),
  #second axis- real value
  sec.axis = sec_axis(~ . +worriedhousingall_2022$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I never worried I would not have enough \n money to cover the cost of my housing",
       subtitle ="UCSD Undergrads never worried about paying for housing",
       y=lab_diff_axis_ug,
       caption=caption_ug)
```

### Worried about Debt/Finances

UCUES has asked "How often during the past academic year have you: Worried about my debt and financial circumstances? since 2014.

The question has six possible responses -from negative - Very Often, Often,Somewhat Often, and occasionally, to two coded as positive: rarely and never.

#### Worried about Finances Change over Time

Overall, UCSD's students are similiar to the UC average over, with only a quarter or so of students never or rarely worrying about finances and debt. Both groups have seen an uptick in student never or rarely worrying in the last wave, to about a third of students.

```{r}
#| label: worried_debt_ug_overtime
#| echo: FALSE
#| warning: FALSE

worried_time<-ucall%>%
  #select question and comparison group - leave all years
  filter(q=="Worried about my debt and financial circumstances"&comp=="All")%>%
  clean_vibe_freq(.)%>%
group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="positive")
 
 g_worried_time<-plot_time(worried_time)+
  #need to tweak scale because otherwise it's too zoomed in
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(.20,.60))+
   labs(title="I never or rarely worried \nabout my debt and financial circumstances",
        caption=caption_ug,
      subtitle = "Percent of respondents never or rarely worried",
         x="UCUES Survey Year")

direct.label(g_worried_time,list(cex=30,  "first.qp"))





```

#### Differences in financial worries between Students

It's notable that while 2022 saw an increase in students never or rarely worrying, stark differences remain by student group, from half of international students to just 22% of Pell Grant recipients.

```{r}
#| label: worried_debt_ug_2022
#| echo: FALSE
#| warning: FALSE
worriedall_2022<-es%>%
filter(q=="Worried about my debt and financial circumstances" & year=="2022"& comp==
"All")%>%
  clean_vibe_freq(.)%>%
collapse_positive_vibe(.)

worried_2022<-es%>%
  filter(q=="Worried about my debt and financial circumstances" & year=="2022")%>%
   clean_vibe_freq(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=worriedall_2022$value)%>%
  create_diff_cols(.)

  
plot_lollipop_div(worried_2022)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.2),
  #second axis- real value
  sec.axis = sec_axis(~ . +worriedall_2022$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I rarely or never worried \n about my debt and financial circumstances",
       subtitle ="UCSD Undergrads never or rarely worried",
       y=lab_diff_axis_ug,
       caption=caption_ug)

```

### Manageable Costs
```{r}
#| label: manageable_ug_overtime
#| echo: FALSE
#| warning: FALSE

manage_ug_time<-ucall%>%
  #select question and comparison group - leave all years
  filter(grepl('manageable', q) & comp=="All")%>%
  clean_vibe_agree(.)%>%
group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="positive")
 
 g_manage_ug_time<-plot_time(manage_ug_time)+

  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(.50,.75))+
   labs(title="Given the grants and scholarships, if any, that you receive, \nthe total cost of attending this university is manageable",
        caption=caption_ug,
      subtitle = "Percent of respondents who agree",
         x="UCUES Survey Year")

direct.label(g_manage_ug_time,list(cex=30,  "first.qp"))
```

```{r}
#| label: manageable_ug_2022
#| echo: FALSE
#| warning: FALSE
manageall_2022<-es%>%
filter(grepl('manageable', q)& comp==
"All")%>%
  clean_vibe_agree(.)%>%
collapse_positive_vibe(.)

manageable_2022<-es%>%
  filter(grepl('manageable', q)& year=="2022")%>%
   clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=manageall_2022$value)%>%
  create_diff_cols(.)

  
plot_lollipop_div(manageable_2022)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.1,.15),
  #second axis- real value
  sec.axis = sec_axis(~ . +manageall_2022$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="Given the grants and scholarships, if any, that you receive, the total cost of attending this university is manageable",
       subtitle ="UCSD Undergrads who somewhat agree, agree or strongly agree",
       y=lab_diff_axis_ug,
       caption=caption_ug)

```





### Additional UCUES Questions related to Affordability.

-   Overall, I feel the education is affordable at my campus
- How satisfied are you with the following aspect of your campus experiences/education? Value of your education for the price you are paying. 
-   How concerned have you been about paying for your undergraduate education up to now?
-   ...Next Year?
-  How often during the past academic year have you: cut down on personal / recreational spending to help pay for college expenses?



## Graduate Students

### Food Security

#### Change over Time

```{r}
#| label: food_g_overtime
#| echo: FALSE
#| warning: FALSE
food_secure_time_grad<-ucall_gses%>%
  #select question and comparison group 
  filter(q=="Food Security")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="positive")

g_foodsecure_time_grad<-plot_time(food_secure_time_grad)+
   labs(title="Food Secure",
        caption=caption_grad,
      subtitle = 'Percent of Food Secure respondents',
         x="GSES Survey Year")

direct.label(g_foodsecure_time_grad,list(cex=30,  "first.qp"))
 
```

#### Food Security Difference by Student Group

```{r}
#| label: food_g_2021
#| echo: FALSE
#| warning: FALSE

foodsecure_all_2021_gses<-gses%>%
filter(q=="Food security"&year=="2021"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
collapse_positive_vibe(.)

foodsecure_2021_gses<-gses%>%
  filter(q=="Food security"&year=="2021")%>%
  clean_vibe_foodsecure(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=foodsecure_all_2021_gses$value)%>%
  create_diff_cols(.)

plot_lollipop_div(foodsecure_2021_gses)+
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.1),
  #second axis- real value
  sec.axis = sec_axis(~.+foodsecure_all_2021_gses$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="Food secure",
       subtitle ="UCSD Graduate Students who are food secure",
       y=lab_diff_axis_grad,
       caption=caption_grad)


```

### Financial hardship has impeded my success in my program.

```{r}
#| label: hardship_g_2021
#| echo: FALSE
#| warning: FALSE

hardship_all_2021_gses<-gses%>%
filter(grepl('hardship', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree_neg(.)%>%
collapse_positive_vibe(.)

hardship_2021_gses<-gses%>%
  filter(q=="Financial hardship has impeded my success in my program."&year=="2021")%>%
  clean_vibe_agree_neg(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=hardship_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_lollipop_div(hardship_2021_gses)  +
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.2),
  #second axis- real value
  sec.axis = sec_axis(~.+hardship_all_2021_gses$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="Financial hardship has impeded my success in my program.",
       subtitle ="Graduate students who somewhat disagree, disagree, strongly disagree",
       y=lab_diff_axis_grad,
       caption=caption_grad)


```



###I've been concerned about money lately.


```{r}
#| label: concerned_g_2021
#| echo: FALSE
#| warning: FALSE

concerned_all_2021_gses<-gses%>%
filter(grepl('concerned', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree_neg(.)%>%
collapse_positive_vibe(.)

concerned_2021_gses<-gses%>%
  filter(q=="I've been concerned about money lately."&year=="2021")%>%
  clean_vibe_agree_neg(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=concerned_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_lollipop_div(concerned_2021_gses)  +
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.2),
  #second axis- real value
  sec.axis = sec_axis(~.+concerned_all_2021_gses$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I've been concerned about money lately.",
       subtitle ="Graduate students who somewhat disagree, disagree, strongly disagree",
       y=lab_diff_axis_grad,
       caption=caption_grad)


```

### I'm confident in my financial situation.

```{r}
#| label: confident_g_2021
#| echo: FALSE
#| warning: FALSE

confident_all_2021_gses<-gses%>%
filter(grepl('confident', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree(.)%>%
collapse_positive_vibe(.)

confident_2021_gses<-gses%>%
  filter(q=="I'm confident in my financial situation."&year=="2021")%>%
  clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=confident_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_lollipop_div(confident_2021_gses)  +
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.2),
  #second axis- real value
  sec.axis = sec_axis(~.+confident_all_2021_gses$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I'm confident in my financial situation.",
       subtitle =subtitle_grad_agree,
       y=lab_diff_axis_grad,
       caption=caption_grad)


```


### I can get by financially without having to cut back on too many of the things that are important to me.

```{r}
#| label: getby_g_2021
#| echo: FALSE
#| warning: FALSE

getby_all_2021_gses<-gses%>%
filter(grepl('get by', q)& year=="2021" & comp=="All")%>%
  clean_vibe_agree(.)%>%
collapse_positive_vibe(.)

getby_2021_gses<-gses%>%
  filter(q=="I can get by financially without having to cut back on too many of the things that are important to me."&year=="2021")%>%
  clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_positive_vibe(.)%>%
  mutate(all=getby_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_lollipop_div(getby_2021_gses)  +
  #first axis- divergence
  scale_y_continuous(labels=scales::percent_format(accuracy=1),limits=c(-.2,.2),
  #second axis- real value
  sec.axis = sec_axis(~.+getby_all_2021_gses$value, labels=scales::label_percent(accuracy=1)))+
  #labels
   labs(title="I can get by financially without having to cut back on too many of the things that are important to me.",
       subtitle =subtitle_grad_agree,
       y=lab_diff_axis_grad,
       caption=caption_grad)


```


