---
title: "Affordable Learning and Financial Support"
editor: source
---

This is a chapter about affordable learning and financial support at University of California San Diego.

```{r}
#| label: run_scripts
#| include: FALSE

source("rscripts/load_packages.R", local = knitr::knit_global())

source("rscripts/functions_clean_data.R" ,local = knitr::knit_global())

source("rscripts/functions_visualize.R" ,local = knitr::knit_global())

```

```{r}
#| label: load_data
#| include: FALSE

#ucues
es<- read_excel("data/ucues_2023.xlsx")%>%
  filter(wg=="Affordable")%>%
  label_groups(.)

ucall<-read_excel("data/ucues_alluc.xlsx")

gses<-read_excel("data/UCGSES.xlsx")%>%
  filter(wg=="Affordable")%>%
  label_groups(.)

ucall_gses<-read_excel("data/gses_alluc.xlsx")

```

```{r}
#| label: viz_captions
#| include: FALSE


#subtitle
subtitle_grad_agree<-"UCSD Graduate Students who somewhat agree, agree, or strongly agree"

#caption
caption_ug<-"Source: UCUES 2022, UC Infocenter"
caption_grad<-"Source: GSES 2021 UC Infocenter"
```

## Undergraduates

This section covers a few questions prioritized by the Affordable Learning and Financial Support collective impact working group.

Please refer to @sec-UCUES for more information on UCUES. For readers affiliated with UCSD and a working Active Directory login, a few longitudinal questions can be explored on [Instiutional Research's Dashboard](https://tableau.ucsd.edu/#/views/UCUES_UC_SanDiego/Affordability?:iid=1)

### Food Security

Student respondents are assigned a food security level: very low food security, low food security, or food secure, based on their responses to six questions.

For more information on this measure, please visit [UCOP Measuring Food Insecurity](https://www.universityofcalifornia.edu/sites/default/files/measuring-food-insecurity.pdf)

#### Food Security over Time

These visuals display the percentage of respondents who are "food insecure" or "very food insecure". For example, in 2022, 42% of UCSD undergraduate respondents were classified as food insecure or very food insecure. This means the remaining 58% of respondents were food secure.

The percent of undergraduates who are food insecure has remained relatively stable over time. UCSD's food insecurity rates are slightly lower or equal to the overall UC system rate for this time period.

```{r}
#| label: food_secure_ug_overtime
#| echo: FALSE
#| warning: FALSE

foodsecure_uccomp<-ucall%>%
  filter(q=="Food Security"&campus=="UC")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="negative")%>%
  ungroup(.)%>%
  mutate(campus="All UC")

food_secure_time<-es%>%
  #select question and comparison group - leave all years
  filter(q=="Food Security"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="negative")%>%
  ungroup(.)%>%
  mutate(campus="UCSD")%>%
  bind_rows(., foodsecure_uccomp)

  
  
food_secure_time_wide<-clean_wide_time(food_secure_time)

  
 plot_time_dumbell(food_secure_time_wide)+
     scale_y_continuous(expand=c(0,0), limits=c(.25, .56), breaks=c(.28,.5), labels=c("Less Food Insecure", "More Food Insecure"), position="right")+
               #labels=scales::percent_format(accuracy=1), )+
             scale_x_discrete(expand=c(0.2,0))+
             #set psuedo grid which looks better with later formatting than the grid line
             geom_segment(data=food_secure_time_wide, aes(x=year, xend=year, y=.25, yend=.5),
                                             color="#b2b2b2", size=0.15)+
    #build diff box
             #spacing of diff bar is determined on y axis
            geom_rect(data=food_secure_time_wide, aes(ymin=.5, ymax=.55, xmin=-Inf, xmax=Inf), fill="grey") +
             geom_text(data=food_secure_time_wide, aes(label=paste0(diff*100, "%"), y=.525, x=year), fontface="bold", size=3)+
            geom_text(data=filter(food_secure_time_wide, year=="2016"),
                                     aes(y=.525, x=year, label="Difference"),
                                     color="black", size=3.1, vjust=-2, fontface="bold") +
     labs(title="Food Insecurity in the UC",
    caption=caption_ug,
         subtitle ='Percent of Food Insecure respondents',
        x="Year")
    

   ggsave("viz/food_insecure_time.png", width=9, height=5, units="in", dpi=300)
  

```

#### Food Security Differences by Group

Focusing on UCSD students in 2022, we find that there are significant differences by student group. In comparison to the all undergraduate proportion of 42% food insecure, the proportion is higher for Pell Grant Recipients (51%), First Generation students (52%),  African American students (55%), Hispanic/Latinx students (52%) and American Indian students (50%).

Overall, there is ~18 percentage point difference between the group with the lowest food insecurity - domestic, White students (37%) and that with the highest, African American students (55%).

```{r}
#| label: food_secure_ug_differences_2022
#| echo: FALSE
#| warning: FALSE
foodsecure_all_2022<-es%>%
filter(q=="Food Security"&year=="2022"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
collapse_negative_vibe(.)

foodsecure_2022<-es%>%
  filter(q=="Food Security"&year=="2022")%>%
  clean_vibe_foodsecure(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=foodsecure_all_2022$value)%>%
  create_diff_cols(.)

  plot_group_comp(foodsecure_2022, cols_neg)%>%
    plot_diff_rect(.,foodsecure_2022,.35,.60, "Less Food Insecurity","More Food Insecurity")+
  #labels
   labs(title="Food Insecurity for UCSD Undergrads, 2022",
       subtitle ="% Food Insecure",
       caption=caption_ug)


 ggsave("viz/food_insecure_2022.png", width=9, height=6, units="in", dpi=300)



```

### Housing Costs

The prioritization poll selected two housing questions from the most recent UCUES 2022 wave- these questions had not been asked previously. There are other housing related questions future versions of this document can add to provide historical context.

The question block prompts: "For the following statements, please indicate whether the statement was never true, sometimes true, or often true for you during the current academic year:"

"I was unable to pay all the cost of my housing on time."

"I was worried I would not have enough money to cover the cost of my housing"

Respondents could select "Never True", "Sometime True" or "Often True." For consistency in selecting negative responses - we focus on "Sometimes True" and "Often True" responses- meaning that respondents were unable to pay and worried about their housing costs.

```{r}
#| label: unable_pay_house_ug_diff_2022
#| echo: FALSE
#| warning: FALSE
payhousingall_2022<-es%>%
filter(q=="I was unable to pay all of the cost of my housing on time."&year=="2022"&comp=="All")%>%
  clean_vibe_true3(.)%>%
collapse_negative_vibe(.)



payhousing_2022<-es%>%
  filter(q=="I was unable to pay all of the cost of my housing on time."&year=="2022")%>%
   clean_vibe_true3(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=payhousingall_2022$value)%>%
  create_diff_cols(.)


plot_group_comp(payhousing_2022, cols_neg)%>%
  plot_diff_rect(.,payhousing_2022, .1,.34, "More Able to Pay"," More Unable to Pay")+
  labs(title="I was unable to pay all the cost of my housing on time",
       subtitle ="UCSD Undergrads who were unable to pay housing costs on time, sometimes or often",
       caption=caption_ug)

 ggsave("viz/payhousing_2022.png", width=9, height=5, units="in", dpi=300)

```

About 20% of students were sometimes or often unable to pay on time. This proportion is slightly lower than the UC average of 23% of students unable to pay (not pictured).

Within UCSD, the proportion of students unable to pay on time ranges from 28% for African American students to 11% for American Indian students, a 17 percentage point difference.

The proportion of students who worry about paying housing costs is higher. Overall 46% worry -similar to the UC average of 45%. However, there is a wide range by student groups at UCSD, from 68% of Pell Grant students worrying, to only 44% of those who don't receive Pell, a difference of over 20 percentage points. The proportions for Hispanic/Latinx and African American students who worry is also high, around 62%.

```{r}
#| label: worried_pay_house_ug_diff_2022
#| echo: FALSE
#| warning: FALSE


worriedhousingall_2022<-es%>%
filter(q=="I worried I would not have enough money to cover the cost of my housing." & year=="2022"& comp==
"All")%>%
  clean_vibe_true3(.)%>%
collapse_negative_vibe(.)

worriedhousing_2022<-es%>%
  filter(q=="I worried I would not have enough money to cover the cost of my housing." & year=="2022")%>%
   clean_vibe_true3(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=worriedhousingall_2022$value)%>%
  create_diff_cols(.)

  plot_group_comp(worriedhousing_2022, cols_neg)%>%
    plot_diff_rect(., worriedhousing_2022, .42,.73,"Less Worried","More Worried")+
    labs(title="I worried I would not have enough money to cover the cost of my housing",
       subtitle ="UCSD Undergrads who were worried, sometimes or often",
       caption=caption_ug)

 ggsave("viz/worried_housing_2022.png", width=9, height=5, units="in", dpi=300)

```

### Worried about Debt/Finances

UCUES has asked "How often during the past academic year have you: Worried about debt and financial circumstances?" since 2014.

The question has six possible responses -three coded negative: Very Often, Often,Somewhat Often, and three as more positive: Occasionally, Rarely, Never. We focus here on the negative responses. 

Overall, UCSD's students are similar to or slightly less worried than the UC average, with the percent of students worried falling for both in 2022.

```{r}
#| label: worried_debt_ug_overtime
#| echo: FALSE
#| warning: FALSE

worried_time<-ucall%>%
  #select question and comparison group - leave all years
  filter(q=="Worried about my debt and financial circumstances"&comp=="All")%>%
  clean_vibe_freq(.)%>%
group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="negative")
 

worried_time_wide<-clean_wide_time(worried_time)

g_worried_time <- plot_time_dumbell(worried_time_wide) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(.62, .82),
    breaks = c(.62, .75),
    labels = c("Less Worried", "More Worried"),
    position = "right"
  ) +
  scale_x_discrete(expand = c(0.2, 0)) +
  
  #build diff box
  #spacing of diff bar is determined on y axis
  geom_rect(data = worried_time_wide,
            aes(
              ymin = .8,
              ymax = .82,
              xmin = -Inf,
              xmax = Inf
            ),
            fill = "grey") +
  geom_text(
    data = worried_time_wide,
    aes(
      label = paste0(diff * 100, "%"),
      y = .81,
      x = year
    ),
    fontface = "bold",
    size = 3
  ) +
  geom_text(
    data = filter(worried_time_wide, year == "2016"),
    aes(y = .81, x = year, label = "Difference"),
    color = "black",
    size = 3.1,
    vjust = -2,
    fontface = "bold"
  ) +
  #set psuedo grid which looks better with later formatting than the grid line
  geom_segment(
    data = worried_time_wide,
    aes(
      x = year,
      xend = year,
      y = .4,
      yend = .6
    ),
    color = "#b2b2b2",
    size = 0.15
  ) +
  labs(title = "UC Comparison: I worried about my debt and financial circumstances",
       caption = caption_ug,
       subtitle = "Percent of respondents worried, somewhat often, often, or very often")


move_layers(g_worried_time, "GeomSegment", position = "bottom")

ggsave(
  "viz/worried_time.png",
  width = 9,
  height = 5,
  units = "in",
  dpi = 300
)




```

It's notable that while 2022 saw a decrease in students worried, stark differences remain by student group,


```{r}
#| label: worried_debt_ug_diff_2022
#| echo: FALSE
#| warning: FALSE
#| 
worried_debt_2022<-es%>%
filter(q=="Worried about my debt and financial circumstances"&year=="2022")%>%
 clean_vibe_freq(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(worried_debt_2022, cols_neg)%>%
  plot_diff_rect(., worried_debt_2022, .47,.83,"Less Worried", "More Worried")+
  labs(title="I worried about my debt and financial circumstances",
         caption=caption_ug,
      subtitle = "Percent of respondents worried, somewhat often, often, or very often")

 ggsave("viz/worried_debt_2022.png", width=9, height=5, units="in", dpi=300)
```

### Manageable Costs

Forty-two percent of students somewhat disagree, disagree, or strongly disagree that the total cost of attending UCSD is manageable, given the the grants and scholarships they receive. Over the past four waves of UCUES, UCSD has had slightly higher proportions of students disagree that it is manageable than the UC average.

```{r}
#| label: manageable_ug_overtime
#| echo: FALSE
#| warning: FALSE

manage_ug_time<-ucall%>%
  #select question and comparison group - leave all years
  filter(grepl('manageable', q) & comp=="All")%>%
  clean_vibe_agree(.)%>%
group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="negative")

 manage_ug_time_wide<-clean_wide_time(manage_ug_time)
 g_manage_time<-plot_time_dumbell(manage_ug_time_wide)+
  scale_y_continuous(expand=c(0,0), limits=c(.37, .52), breaks=c(.37,.46), labels=c("More Manageable", "Less Manageable"), position="right")+
  scale_x_discrete(expand=c(0.2,0))+
    #build diff box
             #spacing of diff bar is determined on y axis
            geom_rect(data=manage_ug_time_wide, aes(ymin=.50, ymax=.52, xmin=-Inf, xmax=Inf), fill="grey") +
             geom_text(data=manage_ug_time_wide, aes(label=paste0(diff*100, "%"), y=.51, x=year), fontface="bold", size=3)+
            geom_text(data=filter(manage_ug_time_wide, year=="2016"),
                                    aes(y=.51, x=year, label="Difference"),
                                    color="black", size=3.1, vjust=-2, fontface="bold") +
   #set psuedo grid which looks better with later formatting than the grid line
             geom_segment(data=manage_ug_time_wide, aes(x=year, xend=year, y=.37, yend=.48),
                                  color="#b2b2b2", size=0.15)+
  labs(title="UC Comparison: Given the grants and scholarships that you receive, \nthe total cost of attending UCSD is manageable",
        caption=caption_ug,
      subtitle = "Percent of respondents who somewhat disagree, disagree, or strongly disagree")
  
  
  move_layers(g_manage_time,"GeomSegment", position="bottom")
  
   ggsave("viz/manageable_time.png", width=9, height=5, units="in", dpi=300)
```

Interestingly, the consistent gaps by student group observed for other affordability questions partially reverse in this question.

Those who do not receive Pell Grants, and not First Generation students have a higher proportion of negative responses. 49% of students not receiving Pell Grants disagree that the cost of UCSD is manageable, compared to 30% of Pell Grant recipients. 

```{r}
#| label: manageable_ug_2022
#| echo: FALSE
#| warning: FALSE
manageall_2022<-es%>%
filter(grepl('manageable', q)& comp==
"All")%>%
  clean_vibe_agree(.)%>%
collapse_negative_vibe(.)

manageable_2022<-es%>%
  filter(grepl('manageable', q)& year=="2022")%>%
   clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=manageall_2022$value)%>%
  create_diff_cols(.)

  
plot_group_comp(manageable_2022, cols_neg)%>%
  plot_diff_rect(., manageable_2022, .29,.54, "More Manageable", "Less Manageable")+
 labs(title="Given the grants and scholarships, if any, that you receive, \nthe total cost of attending UCSD is manageable",
        caption=caption_ug,
      subtitle = "Percent of UCSD students somewhat disagree, disagree, or strongly disagree")

 ggsave("viz/manage_2022.png", width=9, height=5, units="in", dpi=300)
```

### Additional UCUES Questions related to Affordability.

These questions are also available on the 2022 UCUES for further exploration.

-   Overall, I feel the education is affordable at my campus
-   How satisfied are you with the following aspect of your campus experiences/education? Value of your education for the price you are paying.
-   How concerned have you been about paying for your undergraduate education up to now?
-   ...Next Year?
-   How often during the past academic year have you: cut down on personal / recreational spending to help pay for college expenses?

## Graduate Students

There have been fewer consistent questions asked to graduate students over time, and slightly different student groups are available. See data descriptions for more detail. The following draw from the UC Graduate Student Experience Survey in 2021, with the exception of Food Security, which includes data from the 2016 Graduate Student Wellbeing Survey. Please refer to @sec-UCGSES for more on UCGSES.

### Food Security

For the two waves available for Graduate Students, UCSD has similar rates of food insecurity in comparison with the UC average.

```{r}
#| label: food_g_overtime
#| echo: FALSE
#| warning: FALSE
food_secure_time_grad<-ucall_gses%>%
  #select question and comparison group 
  filter(q=="Food Security")%>%
  clean_vibe_foodsecure(.)%>%
  group_by(campus,year, vibe)%>%
  summarize(value=sum(value),
            N=sum(N))%>%
  filter(vibe=="negative")

food_grad_wide<-clean_wide_time(food_secure_time_grad)

plot_time_dumbell(food_grad_wide)+
  scale_y_continuous(expand=c(0,0), limits=c(.18, .30), breaks=c(.37,.46), labels=c("Less Food Insecure", "More Food Insecure"), position="right")+
  scale_x_discrete(expand=c(0.3,0))+
    #build diff box
             #spacing of diff bar is determined on y axis
            geom_rect(data=food_grad_wide, aes(ymin=.28, ymax=.30, xmin=-Inf, xmax=Inf), fill="grey") +
             geom_text(data=food_grad_wide, aes(label=paste0(diff*100, "%"), y=.29, x=year), fontface="bold", size=3)+
            geom_text(data=filter(food_grad_wide, year=="2016"),
                                    aes(y=.29, x=year, label="Difference"),
                                    color="black", size=3.1, vjust=-2, fontface="bold") +
   #set psuedo grid which looks better with later formatting than the grid line
             geom_segment(data=food_grad_wide, aes(x=year, xend=year, y=.2, yend=.28),
                                  color="#b2b2b2", size=0.15)+
  labs(title="UC Comparison: Food Insecurity for Graduate Students",
        caption=caption_grad,
      subtitle = "Percent of respondents food insecure")
  
 ggsave("viz/food_insecure_grad_time.png", width=9, height=5, units="in", dpi=300)
 
```

\*Will add additional student groups to mirror following data.

```{r}
#| label: food_g_2021
#| echo: FALSE
#| warning: FALSE

foodsecure_all_2021_gses<-gses%>%
filter(q=="Food security"&year=="2021"&comp=="All")%>%
  clean_vibe_foodsecure(.)%>%
collapse_negative_vibe(.)

foodsecure_2021_gses<-gses%>%
  filter(q=="Food security"&year=="2021")%>%
  clean_vibe_foodsecure(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=foodsecure_all_2021_gses$value)%>%
  create_diff_cols(.)

plot_group_comp(foodsecure_2021_gses, cols_neg)%>%
  plot_diff_rect(., foodsecure_2021_gses, .15,.41, "Less Food Insecure","More Food Insecure")+
  labs(title="Food Insecurity 2021",
       subtitle ="UCSD Graduate Students who were Food Insecure",
       caption=caption_grad)

ggsave("viz/food_insecure_grad.png", width=9, height=5, units="in", dpi=300)

```

### Financial hardship has impeded my success in my program.

-   add comparison with all UC

Graduate students were asked indicate their agreement with this statement "Financial Hardship has impeded my success in my program." Students who selected "Strongly Agree" "Agree" and "Somewhat Agree" were coded as "negative" responses.

While about 36% of students agreed that financial hardship had impeded their success, that number rises to 52% for Hispanic/Latinx students - and First Generation, LGBT, African American, Doctoral students, and International students also have higher levels of agreement than average.

```{r}
#| label: hardship_g_2021
#| echo: FALSE
#| warning: FALSE

hardship_all_2021_gses<-gses%>%
filter(grepl('hardship', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree_neg(.)%>%
collapse_negative_vibe(.)

hardship_2021_gses<-gses%>%
  filter(q=="Financial hardship has impeded my success in my program."&year=="2021")%>%
  clean_vibe_agree_neg(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=hardship_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_group_comp(hardship_2021_gses, cols_neg)%>%
  plot_diff_rect(., hardship_2021_gses, .25,.57, "Less Hardship","More Hardship")+
labs(title="Financial hardship has impeded \nmy success in my program.",
       subtitle ="UCSD Graduate students who somewhat disagree, disagree, strongly disagree",
       caption=caption_grad)


ggsave("viz/hardship_grad.png", width=9, height=5, units="in", dpi=300)
```

### I've been concerned about money lately.

This question continues the format of the previous, with similiar patterns of differences between groups,

```{r}
#| label: concerned_g_2021
#| echo: FALSE
#| warning: FALSE

concerned_all_2021_gses<-gses%>%
filter(grepl('concerned', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree_neg(.)%>%
collapse_negative_vibe(.)

concerned_2021_gses<-gses%>%
  filter(q=="I've been concerned about money lately."&year=="2021")%>%
  clean_vibe_agree_neg(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=concerned_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_group_comp(concerned_2021_gses, cols_neg)%>%
  plot_diff_rect(., concerned_2021_gses, .56,.89, "Less Concern","More Concern")+
 labs(title="I've been concerned about money lately.",
       subtitle ="Graduate students who somewhat agree, agree, strongly agree",
       caption=caption_grad)


ggsave("viz/concerned_grad.png", width=9, height=5, units="in", dpi=300) 

```

### I'm confident in my financial situation.

For this question, "Strongly Disagree," "Disagree", and "Somewhat Disagree" were considered negative responses.

```{r}
#| label: confident_g_2021
#| echo: FALSE
#| warning: FALSE

confident_all_2021_gses<-gses%>%
filter(grepl('confident', q)&year=="2021"&comp=="All")%>%
  clean_vibe_agree(.)%>%
collapse_negative_vibe(.)

confident_2021_gses<-gses%>%
  filter(q=="I'm confident in my financial situation."&year=="2021")%>%
  clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=confident_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_group_comp(confident_2021_gses, cols_neg)%>%
  plot_diff_rect(., confident_2021_gses, .2,.5,"More Confident", "Less Confident") +
  labs(title="I'm confident in my financial situation.",
       subtitle ="Graduate students who somewhat disagree, disagree, strongly disagree",
       caption=caption_grad)


ggsave("viz/confident_grad.png", width=9, height=5, units="in", dpi=300) 

```


### I can get by financially without having to cut back on too many of the things that are important to me.

For this question, "Strongly Disagree," "Disagree", and "Somewhat Disagree" were considered negative responses.


```{r}
#| label: getby_g_2021
#| echo: FALSE
#| warning: FALSE

getby_all_2021_gses<-gses%>%
filter(grepl('get by', q)& year=="2021" & comp=="All")%>%
  clean_vibe_agree(.)%>%
collapse_negative_vibe(.)

getby_2021_gses<-gses%>%
  filter(q=="I can get by financially without having to cut back on too many of the things that are important to me."&year=="2021")%>%
  clean_vibe_agree(.)%>%
  order_comp_group_all(.)%>%
collapse_negative_vibe(.)%>%
  mutate(all=getby_all_2021_gses$value)%>%
  create_diff_cols(.)%>%
  drop_na(value)

plot_group_comp(getby_2021_gses, cols_neg)%>%
  plot_diff_rect(., getby_2021_gses, .24,.55, "Less Cutting Back", "More Cutting Back")+
   labs(title="I can get by financially without having to cut back on too \nmany of the things that are important to me.",
       subtitle ="Graduate students who somewhat agree, agree, or strongly agree",
       caption=caption_grad)


ggsave("viz/getby_grad.png", width=9, height=5, units="in", dpi=300) 

```
