---
title: "Mentoring Coaching and Advising"
editor: source
---


```{r}
#| label: run_scripts
#| include: FALSE

source("rscripts/load_packages.R", local = knitr::knit_global())

source("rscripts/functions_clean_data.R" ,local = knitr::knit_global())

source("rscripts/functions_visualize.R" ,local = knitr::knit_global())

```

```{r}
#| label: set_labels
#| echo: FALSE
#| warning: FALSE

caption_ug<-"Source: UCUES 2022, UC Infocenter"
caption_grad<-"Source: GSES 2021 UC Infocenter"
```

```{r}
#| label: load_data
#| echo: FALSE
#| warning: FALSE
#ucues
es<- read_excel("data/ucues_2023.xlsx")%>%
  label_groups(.)

gses<-read_excel("data/UCGSES.xlsx")%>%
  label_groups(.)

```
## UCUES

### Faculty know my name

Need to input visual from previous chapter

### Reluctance to ask for academic help when I need it
```{r}
#| label: rel_help
#| echo: FALSE
#| warning: FALSE
rel_help_2022<-es%>%
filter(q=="Reluctance to ask for academic help when I need it"&year=="2022")%>%
 clean_vibe_freq5(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(rel_help_2022, cols_neg)%>%
  plot_diff_rect(., rel_help_2022,.4,.58, "Less of an obstacle", "More of an obstacle")+
  labs(title="How often is reluctance to ask for academic help \nwhen you need it, an obstacle to your academic success?",
         caption=caption_ug,
      subtitle = "Percent of UCSD UG who responded occasionally, frequently, or all the time")
  
 ggsave("viz/mca/reluctant_help_2022.png", width=9, height=5, units="in", dpi=300)
```



### Sought academic help from instructor or tutor when needed

```{r}
#| label: sought_help
#| echo: FALSE
#| warning: FALSE
sought_help_2022<-es%>%
filter(q=="Sought academic help from instructor or tutor when needed"&year=="2022")%>%
 clean_vibe_freq(.)%>%
  #rarely or never is positive
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(sought_help_2022, cols_neg)%>%
  plot_diff_rect(., sought_help_2022,.12,.34, "More helpseeking", "Less helpseeking")+
  labs(title="How often have you sought academic help from \nan instructor or tutor when needed?",
         caption=caption_ug,
      subtitle = "Percent of UCSD UG who responded rarely or never")
 ggsave("viz/mca/sought_help_2022.png", width=9, height=5, units="in", dpi=300)
```

### Studied as a group with classmates outside of class

```{r}
#| label: study_group
#| echo: FALSE
#| warning: FALSE

study_group_2022<-es%>%
filter(q=="Studied as a group with classmates outside of class"&year=="2022")%>%
 clean_vibe_freq(.)%>%
  #rarely or never is "positive" in clean vibe function - so we want to collapse those
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(study_group_2022, cols_neg)%>%
  plot_diff_rect(., study_group_2022,.22,.42, "More group study", "Less group study")+
  labs(title="How often have you: studied as a group \nwith classmates outside of class ?",
         caption=caption_ug,
      subtitle = "Percent of UCSD UG who responded rarely or never")
 ggsave("viz/mca/group_study_2022.png", width=9, height=5, units="in", dpi=300)

```

### Helped a classmate better understand the course material when studying together

```{r}
#| label: help_classmate
#| echo: FALSE
#| warning: FALSE

help_class_2022<-es%>%
filter(q=="Helped a classmate better understand the course material when studying together"&year=="2022")%>%
 clean_vibe_freq(.)%>%
  #rarely or never is "positive" in clean vibe function - so we want to collapse those
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(help_class_2022, cols_neg)%>%
  plot_diff_rect(., help_class_2022,.12,.36, "More helping", "Less helping")+
  labs(title="How often have you: Helped a classmate \nbetter understand the course material when studying together",
         caption=caption_ug,
      subtitle = "Percent of UCSD UG who responded rarely or never")
 ggsave("viz/mca/help_class_2022.png", width=9, height=5, units="in", dpi=300)


```


### How many professors do you know well enough to ask for a letter of recommendation?
```{r}
#| label: num_prof
#| echo: FALSE
#| warning: FALSE



num_prof_2022<-es%>%
filter(q=="How many professors do you know well enough to ask for a letter of recommendation?"&year=="2022")%>%
  mutate(repsonse=as_factor(response))%>%
    mutate(vibe=fct_collapse(response,
                        negative=c("0 Professor"),
                        positive=c("4 Professors","3 Professors", "2 Professors","1 Professor")))%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(num_prof_2022, cols_neg)%>%
  plot_diff_rect(., num_prof_2022,.38,.60, "Fewer know 0", "More know 0")+
  labs(title="I know 0 professors well enough to ask for a letter of recommendation",
         caption=caption_ug,
      subtitle = "Percent of UCSD UG who know 0 professors well enough")
 ggsave("viz/mca/num_prof_2022.png", width=9, height=5, units="in", dpi=300)
```

## UCGSES

### Opportunities to form research/academic mentorship relationships with faculty members
```{r}
#| label: opp_research_ment
#| echo: FALSE
#| warning: FALSE
opp_research_ment_2021<-gses%>%
  filter(q=="Opportunities to form research/academic mentorship relationships with faculty members")%>%
  clean_vibe_rate(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(opp_research_ment_2021, cols_neg)%>%
  plot_diff_rect(., opp_research_ment_2021,.18,.4, "Fewer Low Ratings", "More Low Ratings")+
  labs(title="Please Rate: Opportunities to form research/academic \nmentorship relationships with faculty members",
         caption=caption_grad,
      subtitle = "Percent of UCSD Graduate Students who respond Very Poor, Poor or Fair")
 ggsave("viz/mca/opp_research_ment2021.png", width=9, height=5, units="in", dpi=300)


```


### Opportunities to form research/academic mentorship relationships with faculty members
```{r}
#| label: opp_ment
#| echo: FALSE
#| warning: FALSE
opp_ment_2021<-gses%>%
  filter(q=="Opportunities to form non-research/academic mentorship relationships with faculty members in your program")%>%
  clean_vibe_rate(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(opp_ment_2021, cols_neg)%>%
  plot_diff_rect(., opp_ment_2021,.3,.52, "Fewer Low Ratings", "More Low Ratings")+
  labs(title="Please Rate: Opportunities to form non-research/academic \nmentorship relationships with faculty members",
    caption=caption_grad,
      subtitle = "Percent of UCSD Graduate Students who respond Very Poor, Poor or Fair")
 ggsave("viz/mca/opp_ment2021.png", width=9, height=5, units="in", dpi=300)


```

### The career support I receive in my program

```{r}
#| label: career_supp
#| echo: FALSE
#| warning: FALSE

career_sup_2021<-gses%>%
  filter(q=="the career support I receive in my program")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(career_sup_2021, cols_neg)%>%
  plot_diff_rect(., career_sup_2021,.15,.35, "Fewer Dissatisfied", "More Dissatisfied")+
  labs(title="I'm satisfied with the career support I recieve in my program",
    caption=caption_grad,
      subtitle = "Percent of UCSD Graduate Students who Strongly Disagree, Disagree, Somewhat Disagree")
 ggsave("viz/mca/career_supp2021.png", width=9, height=5, units="in", dpi=300)


```

### The mentorship and advising I receive in my program

```{r}
#| label: grad_mentorship
#| echo: FALSE
#| warning: FALSE

mentor_gen_2021<-gses%>%
  filter(q=="the mentorship and advising I receive in my program")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(mentor_gen_2021, cols_neg)%>%
  plot_diff_rect(., mentor_gen_2021,.12,.3, "Fewer Dissatisfied", "More Dissatisfied")+
  labs(title="I'm satisfied with the mentorship and advising I receive in my program",
    caption=caption_grad,
      subtitle = "Percent of UCSD Graduate Students who Strongly Disagree, Disagree, Somewhat Disagree")
 ggsave("viz/mca/ment_gen2021.png", width=9, height=5, units="in", dpi=300)


```

### The support I received regarding my thesis/dissertation research

```{r}
#| label: support_thesis
#| echo: FALSE
#| warning: FALSE

diss_supp_2021<-gses%>%
  filter(q=="the support I received regarding my thesis/dissertation research")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(diss_supp_2021, cols_neg)%>%
  plot_diff_rect(., diss_supp_2021,.09,.28, "Fewer Dissatisfied", "More Dissatisfied")+
  labs(title="I'm satisfied with the support I received regarding \nmy thesis/dissertation research",
    caption=caption_grad,
      subtitle = "Percent of UCSD Graduate Students who Strongly Disagree, Disagree, Somewhat Disagree")
 ggsave("viz/mca/diss_supp_20212021.png", width=9, height=5, units="in", dpi=300)


```
## Response Rates

#### UCUES

```{r}
#| label: rr_ug_data
#| echo: FALSE
#| warning: FALSE
ug_rr<-es%>%
  filter(q=="Had a class in which the professor knew or learned your name"|q=="Reluctance to ask for academic help when I need it"|q=="Sought academic help from instructor or tutor when needed"|q=="How many professors do you know well enough to ask for a letter of recommendation?")%>%
  filter(year=="2022")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_ug_gt
#| echo: FALSE
#| warning: FALSE
gt_ug_rr<- ug_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCUES 2022")%>%
  tab_spanner(
    label = "UCUES Questions",
    columns=3:6)
  

gtsave(gt_ug_rr, "mca_rr_ug.png", path = "viz/mca")
```

### UCGSES
```{r}
#| label: rr_grad_data
#| echo: FALSE
#| warning: FALSE
grad_rr<-gses%>%
  filter(q=="Opportunities to form research/academic mentorship relationships with faculty members"|
           q=="Opportunities to form non-research/academic mentorship relationships with faculty members in your program"| 
           q=="the career support I receive in my program"|
           q=="the career support I receive in my program"|
           q=="the mentorship and advising I receive in my program"|
           q=="the support I received regarding my thesis/dissertation research")%>%
  filter(year=="2021")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_grad_gt
#| echo: FALSE
#| warning: FALSE
gt_grad_rr<- grad_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
 
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCGSES 2021")%>%
  tab_spanner(
    label = "UCGSES Questions",
    columns=3:7)

  


gtsave(gt_grad_rr, "mca_rr_grad.png", path = "viz/mca")

```


