---
title: "Sense of Belonging"
editor: source
 html:
    fig-width: 8
    fig-height: 5.75
    
---
This is a chapter about Sense of Belonging at University of California San Diego. @sec-sob-undergrads details undergraduates' experiences, using information from the UCUES survey. @sec-sob-grads explores graduate student experiences using the UCGSES survey. Finally, an appendix @sec-sob-rr provides the number of respondents for each survey question by student group. 

Readers should note that the number of American Indian student respondents is low,about 36 respondents for the UCUES questions and too few to report for UCGSES. While this low response number is in line with the low number of American Indian students at UCSD, the low number means one or two students' responses can change the overall takeaway for any given question. You may see high variability in responses for this group as a result. 

This chapter continues be updated and edited with additional interpretation and student groups - most recent updates 3/23/23. 
```{r}
#| label: run_scripts
#| include: FALSE

source("rscripts/load_packages.R", local = knitr::knit_global())

source("rscripts/functions_clean_data.R" ,local = knitr::knit_global())

source("rscripts/functions_visualize.R" ,local = knitr::knit_global())
```

```{r}
#| label: set_labels
#| echo: FALSE
#| warning: FALSE

caption_ug<-"Source: UCUES 2022, UC Infocenter"
caption_grad<-"Source: GSES 2021 UC Infocenter"
```

```{r}
#| label: load_data
#| echo: FALSE
#| warning: FALSE
#ucues
es<- read_excel("data/ucues_2023.xlsx")%>%
  label_groups(.)

gses<-read_excel("data/UCGSES.xlsx")%>%
  label_groups(.)

```

## Undergraduates (UCUES 2022) {#sec-sob-undergrads}

### I feel I belong at this University

```{r}
#| label: belong_2022
#| echo: FALSE
#| warning: FALSE
belong_2022<-es%>%
filter(q=="I feel that I belong at this university"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(belong_2022, cols_neg)%>%
  plot_diff_rect(., belong_2022,.14,.36, "Fewer Feeling \nDon't Belong", "More Feeling \nDon't Belong")+
  labs(title="I feel that I belong at this university",
         caption=caption_ug,
      subtitle = "UCSD UG who somewhat disagree, disagree, or strongly disagree")
  
 ggsave("viz/sob/belong_2022.png", width=9, height=5, units="in", dpi=300)
```


### I feel valued as an individual

```{r}
#| label: valued_2022
#| echo: FALSE
#| warning: FALSE

valued_2022<-es%>%
filter(q=="I feel valued as an individual at this institution"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(valued_2022, cols_neg)%>%
  plot_diff_rect(., valued_2022,.13,.44, "Less Feeling \nNot Valued", "More Feeling \nNot Valued")+
  labs(title="I feel valued as an individual at this institution",
         caption=caption_ug,
      subtitle = "UCSD UG who somewhat disagree, disagree, or strongly disagree")
  
 ggsave("viz/sob/valued_2022.png", width=9, height=5, units="in", dpi=300)
```


### Overall Social Experience  
```{r}
#| label: social_ex
#| echo: FALSE
#| warning: FALSE
overall_soc_2022<-es%>%
filter(q=="Overall social experience"&year=="2022")%>%
 clean_vibe_satisfied(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(overall_soc_2022, cols_neg)%>%
  plot_diff_rect(., overall_soc_2022,.28,.47, "Less Dissatisfied", "More Dissatisfied")+
  labs(title="Please Rate: Overall Social Experience",
         caption=caption_ug,
      subtitle = "UCSD UG who are somewhat dissatisfied, dissatisfied, or very dissatisfied")
  
 ggsave("viz/sob/overall_soc_2022.png", width=9, height=5, units="in", dpi=300)
```
### UC campus is Welcoming
```{r}
#| label: welcome_2022
#| echo: FALSE
#| warning: FALSE
welcome_2022<-es%>%
filter(q=="UC [CAMPUS] is a welcoming campus"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(welcome_2022, cols_neg)%>%
  plot_diff_rect(., welcome_2022,.06,.31, "Less Disagree", "More Disagree")+
  labs(title="UCSD is a welcoming campus",
         caption=caption_ug,
      subtitle = "UCSD UG who somewhat disagree, disagree, or strongly disagree")
  
 ggsave("viz/sob/welcome_2022.png", width=9, height=5, units="in", dpi=300)
```


```{r}
#| label: climate_edi_2022
#| echo: FALSE
#| warning: FALSE
#| 
climate_edi_2022<-es%>%
filter(q=="Overall, I feel comfortable with the climate for diversity and inclusiveness at this university"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(climate_edi_2022, cols_neg)%>%
  plot_diff_rect(., climate_edi_2022,.06,.45, "Less Disagree", "More Disagree")+
  labs(title="Overall, I feel comfortable with the climate for \ndiversity and inclusiveness at this university",
         caption=caption_ug,
      subtitle = "UCSD UG who somewhat disagree, disagree, or strongly disagree")
  
 ggsave("viz/sob/climate_edi_2022.png", width=9, height=5, units="in", dpi=300)



```

### Channels for addressing discrimination
```{r}
#| label: channel_dis
#| echo: FALSE
#| warning: FALSE


channel_diss_2022<-es%>%
filter(q=="Channels for expressing discrimination or harassment complaints are readily available to students at UC"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(channel_diss_2022, cols_neg)%>%
  plot_diff_rect(., channel_diss_2022,.11,.47, "Less Disagree", "More Disagree")+
  labs(title="Channels for expressing discrimination or harassment complaints are readily available to students at UC",
         caption=caption_ug,
      subtitle = "UCSD UG who somewhat disagree, disagree, or strongly disagree")
  
 ggsave("viz/sob/channel_diss_2022.png", width=9, height=5, units="in", dpi=300)




```


## Graduate Students (UCGSES) {#sec-sob-grads}

### I feel included by my peers
```{r}
#| label: include_peers
#| echo: FALSE
#| warning: FALSE


include_peers_2021<-gses%>%
  filter(q=="I feel included by... my peers")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_peers_2021, cols_neg)%>%
  plot_diff_rect(., include_peers_2021,.05,.17, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by my peers",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_2021.png", width=9, height=5, units="in", dpi=300)
```

### I feel included by... the administration and staff"

```{r}
#| label: include_admin
#| echo: FALSE
#| warning: FALSE
 
include_admin_2021<-gses%>%
  filter(q=="I feel included by... the administration and staff")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_admin_2021, cols_neg)%>%
  plot_diff_rect(., include_admin_2021,.11,.24, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by the administration and staff",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_admin_2021.png", width=9, height=5, units="in", dpi=300)
```
 ###I feel included by the faculty

```{r}
#| label: include_fac
#| echo: FALSE
#| warning: FALSE
 
include_fac_2021<-gses%>%
  filter(q=="I feel included by... the faculty")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_fac_2021, cols_neg)%>%
  plot_diff_rect(., include_fac_2021,.10,.24, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by the faculty",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_fac_2021.png", width=9, height=5, units="in", dpi=300)
```

## Response Rates {#sec-sob-rr}

### Undergraduates(UCUES)

```{r}
#| label: rr_ug_data
#| echo: FALSE
#| warning: FALSE
ug_rr<-es%>%
  filter(q=="I feel that I belong at this university"|
           q=="I feel included by... the administration and staff"|
           q=="UC [CAMPUS] is a welcoming campus"|
           q=="Channels for expressing discrimination or harassment complaints are readily available to students at UC"| 
           q=="Overall social experience")%>%
  filter(year=="2022")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_ug_gt
#| echo: FALSE
#| warning: FALSE
gt_ug_rr<- ug_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCUES 2022")%>%
  tab_spanner(
    label = "UCUES Questions",
    columns=3:6)
  

gtsave(gt_ug_rr, "sob_rr_ug.png", path = "viz/sob")
```

### Graduates Students(UCGSES)

```{r}
#| label: rr_grad_data
#| echo: FALSE
#| warning: FALSE
grad_rr<-gses%>%
  filter(q=="I feel included by... my peers"|
           q=="I feel included by... the faculty"| 
           q=="I feel included by... the administration and staff")%>%
  filter(year=="2021")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_grad_gt
#| echo: FALSE
#| warning: FALSE
gt_grad_rr<- grad_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
 
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCGSES 2021")%>%
  tab_spanner(
    label = "UCGSES Questions",
    columns=3:5)

  


gtsave(gt_grad_rr, "sob_rr_grad.png", path = "viz/sob")

```



