---
title: "Sense of Belonging"
editor: source
format:
  html:
    fig-width: 8
    fig-height: 5.75
    
---
This is a chapter about Sense of Belonging at University of California San Diego. @sec-sob-undergrads details undergraduates' experiences, using information from the UCUES survey. @sec-sob-grads explores graduate student experiences using the UCGSES survey. Finally, an appendix @sec-sob-rr provides the number of respondents for each survey question by student group. 

Readers should note that the number of American Indian student respondents is low,about 36 respondents for the UCUES questions and too few to report for UCGSES. While this low response number is in line with the low number of American Indian students at UCSD, the low number means one or two students' responses can change the overall takeaway for any given question. You may see high variability in responses for this group as a result. 

This chapter continues be updated and edited with additional interpretation and student groups - most recent updates 3/23/23. 
```{r}
#| label: run_scripts
#| include: FALSE

source("rscripts/load_packages.R", local = knitr::knit_global())

source("rscripts/functions_clean_data.R" ,local = knitr::knit_global())

source("rscripts/functions_visualize.R" ,local = knitr::knit_global())
```

```{r}
#| label: set_labels
#| echo: FALSE
#| warning: FALSE

caption_ug<-"Source: UCUES 2022, UC Infocenter"
caption_grad<-"Source: GSES 2021 UC Infocenter"
```

```{r}
#| label: load_data
#| echo: FALSE
#| warning: FALSE
#ucues
es<- read_excel("data/ucues_2023.xlsx")%>%
  label_groups(.)

gses<-read_excel("data/UCGSES.xlsx")%>%
  label_groups(.)

#slightly different download pattern - looking at https://www.universityofcalifornia.edu/about-us/information-center/ucues-longitudinal you can download already aggregated negative and positve responses
ucall_longit<-read_excel("data/ucues_longit_agg.xlsx")

```

## Undergraduates (UCUES 2022) {#sec-sob-undergrads}

This section draws from recent waves of the University of California Undergraduate Experience Survey. 
Please refer to @sec-UCUES for more information on UCUES. For readers affiliated with UCSD and a working Active Directory login, a few longitudinal questions can be explored on [Institutional Research's Dashboard](https://tableau.ucsd.edu/#/views/UCUES_UC_SanDiego/AcademicEngagement?:iid=1)


### Sense of Belonging at the University Level

UC San Diego undergraduates have consistently reported feeling a lower sense of belonging at their university than the UC system average. However, 2022 brought an increase to UCSD -with 80% of students at least somewhat agreeing they feel they belong at UCSD and the lowest difference between the UC average. 
 

```{r}
#| label: ug_belong_overtime
#| echo: FALSE
#| warning: FALSE

belong_uccomp<-ucall_longit%>%
  filter(q=="I feel that I belong at this university")%>%
   filter(vibe=="positive")%>%
  group_by(campus,year)%>%
  ungroup(.)


  
  
belong_uccomp_wide<-clean_wide_time(belong_uccomp)


 

 plot_time_dumbell(belong_uccomp_wide)+
     scale_y_continuous(expand=c(0,0), limits=c(.70, .89), breaks=c(.72,.79), labels=c("Less Belonging", "More Belonging"), position="right")+
               #labels=scales::percent_format(accuracy=1), )+
             scale_x_discrete(expand=c(0.2,0))+
             #set psuedo grid which looks better with later formatting than the grid line
             geom_segment(data=belong_uccomp_wide, aes(x=year, xend=year, y=.25, yend=.5),
                                             color="#b2b2b2", size=0.15)+
    #build diff box
             #spacing of diff bar is determined on y axis
            geom_rect(data=belong_uccomp_wide, aes(ymin=.86, ymax=.88, xmin=-Inf, xmax=Inf), fill="grey") +
             geom_text(data=belong_uccomp_wide, aes(label=paste0(diff*100, "%"), y=.87, x=year), fontface="bold", size=3)+
            geom_text(data=filter(belong_uccomp_wide, year=="2006"),
                                     aes(y=.87, x=year, label="Difference"),
                                     color="black", size=3.1, vjust=-2, fontface="bold") +
     labs(title="Positive Sense of Belonging in the UC",
        x="Year")
    

   ggsave("viz/sob/belong_time.png", width=9, height=5.5, units="in", dpi=300)
  

```


#### I feel I belong at this University
Substantial differences exist between student groups within UCSD, with Black undergraduates reporting the lowest rates of feeling like they belong, are valued, and respected on campus. This summary is far from the first to document this disparity. The [Black Academic Excellence Initiative](https://diversity.ucsd.edu/initiatives/black-academic-excellence/index.html) provides a more holistic review and summary of current programmatic efforts to change how the university supports Black students.
```{r}
#| label: belong_2022
#| echo: FALSE
#| warning: FALSE
belong_2022<-es%>%
filter(q=="I feel that I belong at this university"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(belong_2022, cols_pos)%>%
  plot_diff_rect(., belong_2022,.67,.92, "Fewer Feel \nBelonging", "")+
  labs(title="I feel that I belong at this university",
         caption=caption_ug,
      subtitle = "UCSD UG who at least somewhat agree")
  
 ggsave("viz/sob/belong_2022.png", width=9, height=5, units="in", dpi=300)
```


#### I feel valued as an individual at this institution

```{r}
#| label: valued_2022
#| echo: FALSE
#| warning: FALSE

valued_2022<-es%>%
filter(q=="I feel valued as an individual at this institution"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(valued_2022, cols_pos)%>%
  plot_diff_rect(., valued_2022,.61,.90, "Less Feeling \nValued", "More Feeling \nValued")+
  labs(title="I feel valued as an individual at this institution",
         caption=caption_ug,
      subtitle = "UCSD UG who at least somewhat agree")
  
 ggsave("viz/sob/valued_2022.png", width=9, height=5, units="in", dpi=300)
```



#### UCSD is a welcoming campus
```{r}
#| label: welcome_2022
#| echo: FALSE
#| warning: FALSE
welcome_2022<-es%>%
filter(q=="UC [CAMPUS] is a welcoming campus"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(welcome_2022, cols_pos)%>%
  plot_diff_rect(., welcome_2022,.76,.99, "Less Agree", "")+
  labs(title="UCSD is a welcoming campus",
         caption=caption_ug,
      subtitle = "UCSD UG who at least somewhat agree")
  
 ggsave("viz/sob/welcome_2022.png", width=9, height=5, units="in", dpi=300)
```

While overall about 86% of UC undergraduates feel comfortable with the climate for diversity and inclusion on campus, and about 80% feel that channels are readily available to address discrimination, these proportions are much lower for African American students, and slightly lower than average for Hispanic/Latinx and American Indian Students.  


#### Comfort with the climate for diversity and inclusiveness

```{r}
#| label: climate_edi_2022
#| echo: FALSE
#| warning: FALSE
#| 
climate_edi_2022<-es%>%
filter(q=="Overall, I feel comfortable with the climate for diversity and inclusiveness at this university"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(climate_edi_2022, cols_pos)%>%
  plot_diff_rect(., climate_edi_2022,.58,.98, "More Disagree", "")+
  labs(title="Overall, I feel comfortable with the climate for \ndiversity and inclusiveness at this university",
         caption=caption_ug,
      subtitle = "UCSD UG who at least somewhat agree")
  
 ggsave("viz/sob/climate_edi_2022.png", width=9, height=5, units="in", dpi=300)



```

#### Availability of channels for addressing discrimination
```{r}
#| label: channel_dis
#| echo: FALSE
#| warning: FALSE


channel_diss_2022<-es%>%
filter(q=="Channels for expressing discrimination or harassment complaints are readily available to students at UC"&year=="2022")%>%
 clean_vibe_agree(.)%>%
  collapse_positive_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(channel_diss_2022, cols_pos)%>%
  plot_diff_rect(., channel_diss_2022,.58,.93, "Less Agree", "")+
  labs(title="Channels for expressing discrimination or harassment complaints are \nreadily available to students at UCSD",
         caption=caption_ug,
      subtitle = "UCSD UG who at least somewhat agreee")
  
 ggsave("viz/sob/channel_diss_2022.png", width=9, height=5, units="in", dpi=300)




```
#### To be visualized

* Top campus administrators are genuinely committed to promoting respect for and understanding of group differences at  UC
* Students of my [… ] are respected on this campus
  + socio-economic status
  + sexual orientation
  + religious beliefs
  + race/ethnicity
  + Political beliefs
  + Gender
  + Immigrant background
  + Disability of condition




### Sense of Belonging within Classes, Majors, with Faculty

This section is under development - available UCUES questions include:

* Overall, I feel comfortable with the climate for diversity and inclusion in my major
* ... in my classes
* In this academic year, I have heard about teaching faculty or instructors express negative views about…
  + Genders
  + Immigrant backgrounds
  + Learning of psychological disabilities
  + Physical disabilities
  + Political affiliation, opinion, or beliefs
  + Races or ethnicities
  + Religions
  + Sexual orientations
  + Social Classes

#### Student Social Life, Involvement, and Belonging

#### Overall Social Experience  
```{r}
#| label: social_ex
#| echo: FALSE
#| warning: FALSE
overall_soc_2022<-es%>%
filter(q=="Overall social experience"&year=="2022")%>%
 clean_vibe_satisfied(.)%>%
  collapse_negative_vibe(.)%>%
   mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  #low sample size
  filter(group_label!="All")

plot_group_comp(overall_soc_2022, cols_neg)%>%
  plot_diff_rect(., overall_soc_2022,.28,.47, "Less Dissatisfied", "More Dissatisfied")+
  labs(title="Please Rate: Overall Social Experience",
         caption=caption_ug,
      subtitle = "UCSD UG who are somewhat dissatisfied, dissatisfied, or very dissatisfied")
  
 ggsave("viz/sob/overall_soc_2022.png", width=9, height=5, units="in", dpi=300)
```

#### Engagement in High Impact Practices
While not strictly about Sense of Belonging, a summary of student involvement in service learning, research, creative projects, and internships can be found in @sec-iie-hip in the Academic and Learning Experiences chapter

#### To be visualized
* Have you been or are you currently involved in a student organization?
* Are you or have you been an officer of a student organization?
* Time Allocation: social activities etc.


## Graduate Students (UCGSES) {#sec-sob-grads}

The student groups available in publicly available data differ slightly for graduate students. The following draw from the UC Graduate Student Experience Survey in 2021.  Please refer to @sec-UCGSES for more on UCGSES. Please note that there are a very small number of "Professional Practice" (MD, Pharm.D.) student respondents (~31) on these questions, which skew more positive than the graduate student body average. The response count for African American graduate students is also small (18).  Interpret these trends with additional caution. 

### I feel included by my peers
```{r}
#| label: include_peers
#| echo: FALSE
#| warning: FALSE


include_peers_2021<-gses%>%
  filter(q=="I feel included by... my peers")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_peers_2021, cols_neg)%>%
  plot_diff_rect(., include_peers_2021,.0,.18, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by my peers",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_2021.png", width=9, height=5, units="in", dpi=300)
```

### I feel included by... the administration and staff"

```{r}
#| label: include_admin
#| echo: FALSE
#| warning: FALSE
 
include_admin_2021<-gses%>%
  filter(q=="I feel included by... the administration and staff")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_admin_2021, cols_neg)%>%
  plot_diff_rect(., include_admin_2021,0,.24, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by the administration and staff",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_admin_2021.png", width=9, height=5, units="in", dpi=300)
```
 ###I feel included by the faculty

```{r}
#| label: include_fac
#| echo: FALSE
#| warning: FALSE
 
include_fac_2021<-gses%>%
  filter(q=="I feel included by... the faculty")%>%
  clean_vibe_agree(.)%>%
  collapse_negative_vibe(.)%>%
    mutate(all=filter(., comp=="All")%$% value)%>%
  create_diff_cols(.)%>%
  filter(group_label!="All")

plot_group_comp(include_fac_2021, cols_neg)%>%
  plot_diff_rect(., include_fac_2021,.08,.24, "Fewer Disagree", "More Disagree")+
  labs(title="I feel included by the faculty",
         caption=caption_grad,
      subtitle = "UCSD Graduate Students who respond somewhat disagree, disagree, or strongly disagree")
 ggsave("viz/sob/included_fac_2021.png", width=9, height=5, units="in", dpi=300)
```

## Response Rates {#sec-sob-rr}

### Undergraduates(UCUES)

```{r}
#| label: rr_ug_data
#| echo: FALSE
#| warning: FALSE
ug_rr<-es%>%
  filter(q=="I feel that I belong at this university"|
           q=="I feel included by... the administration and staff"|
           q=="UC [CAMPUS] is a welcoming campus"|
           q=="Channels for expressing discrimination or harassment complaints are readily available to students at UC"| 
           q=="Overall social experience")%>%
  filter(year=="2022")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_ug_gt
#| echo: FALSE
#| warning: FALSE
gt_ug_rr<- ug_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCUES 2022")%>%
  tab_spanner(
    label = "UCUES Questions",
    columns=3:6)
  

gtsave(gt_ug_rr, "sob_rr_ug.png", path = "viz/sob")
```

### Graduates Students(UCGSES)

```{r}
#| label: rr_grad_data
#| echo: FALSE
#| warning: FALSE
grad_rr<-gses%>%
  filter(q=="I feel included by... my peers"|
           q=="I feel included by... the faculty"| 
           q=="I feel included by... the administration and staff")%>%
  filter(year=="2021")%>%
  group_by(q,comp,group_label)%>%
  summarise(N=sum(N, na.rm=T))
```

```{r}
#| label: rr_grad_gt
#| echo: FALSE
#| warning: FALSE
gt_grad_rr<- grad_rr%>%
  pivot_wider(names_from=q, values_from = N)%>%
 
  gt(., rowname_col="group_label", groupname_col="comp")%>%
  cols_align(
    align="left",
    columns=group_label
  )%>%
  tab_header(title="UCSD Student Response Counts",
             subtitle="UCGSES 2021")%>%
  tab_spanner(
    label = "UCGSES Questions",
    columns=3:5)

  


gtsave(gt_grad_rr, "sob_rr_grad.png", path = "viz/sob")

```



